Forecasting the 2024 U.S. Presidential Election Results: Linear Regression, Bayesian and Spline-Fit Gaussian Model by Using Pollsters' Data From Five Thirty Eight

### 2024-11-04 paper/paper.qmd
introduction and abstract updated

---
title: "Forecasting the 2024 U.S. Presidential Election Results: Linear Regression, Bayesian and Spline-Fit Gaussian Model by Using Pollsters' Data From Five Thirty Eight"
subtitle: "Contrasting Harris's Lead: Strong in Linear Regression and Bayesian Model, Modest in Spline-Fit Gaussian Model"
author: 
  - Ziqi Zhu
  - Yuanchen Miao
  - Claire Chang
thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
date: today
date-format: long
abstract: "we use FiveThirtyEight’s 2024 U.S. presidential poll-of-polls data to analyze support trends for the leading candidates, Kamala Harris and Donald Trump. By focusing on key reliability metrics—such as sample size, numeric scores, poll scores, and state-level factors—we fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model. Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model. These results offer insight into the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

@@

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election Results: Linear Regression, Bayesian and Spline-Fit Gaussian Model by Using Pollsters' Data From Five Thirty Eight"
subtitle: "Contrasting Harris's Lead: Strong in Linear Regression and Bayesian Model, Modest in Spline-Fit Gaussian Model"
author: 
  - Ziqi Zhu
  - Yuanchen Miao
  - Claire Chang
thanks: "Code and data are available at: https://github.com/zzq20010617/2024_USelection_prediction"
date: today
date-format: long
abstract: "we use FiveThirtyEight’s 2024 U.S. presidential poll-of-polls data from [date] to [date] to analyze support trends for the leading candidates, Kamala Harris and Donald Trump. By focusing on key reliability metrics, such as sample size, numeric scores, poll scores, and state-level factors. We fit the data into three models: a linear regression model, a Bayesian model, and a spline-fit Gaussian model. Our findings reveal a strong lead for Harris in both the linear regression and Bayesian models, with a more modest lead in the spline-fit Gaussian model. These results offer insight into the potential trajectory of the election, helping the public anticipate the policy directions that might follow a Harris administration."
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

@@

Nowadays, U.S. presidential elections have been marked by intense public interest and polarization. The race between Donald Trump and Kamala Harris, both of them are having a significant leads over other candidates, is currently the most closely watched contest. Polling data is now playing a very important role in showing the public perceptions and support rate to each candidates. Analyzing polling data with statistical models, we are able to gain a valuable insight into candidate support rate across different states and parties. These insights could help the public to predict the shifts in national, economic and social directions. Since the stances hold by Trump and Harris on key issues like healthcare, immigration, and the international relations are in distinct positions, analyzing on polling data to predict the winner of the 2024 U.S. presidential election is been crucial. This paper is contributing to this understanding using the latest polling data and statistical models to capture the trends in support rates of these two candidates.

In this paper, polling data of 2024 U.S. presidential election from FiveThirtyEight [@polls], focusing specifically on the support rate of the leading candidates, Kamala Harris and Donald Trump, is analyzed. The polling data from FiveThirtyEight [@polls] aggregates polls from various sources, thus, to ensure the reliability on the forecast result and analysis, we reduced the variables of the data to only retain variables that reflect polling reliability, such as sample size, numeric scores and poll scores. Additionally, the state-level data is kept as a factor influencing support rates, acknowledging the geographic variations in voter preferences.

The analysis employs three statistical models: linear regression model, Bayesian model and spline-fit Gaussian model. The findings indicates that Kamala Harris maintains a strong lead in both linear regression model and Bayesian model, while her leads becomes modest when comparing the trend of support rate of Donald Trump in spline-fit Gaussian model. Comparing to Donald Trump, Kamala Harris has a [pct] lead in linear regression model, [pct] in Bayesian model and only [pct] lead in spline-fit Gaussian model.

These results are significant as they provide insights into the evolving dynamics of the election and how different statistical models and approaches could yield varying interpretations of candidate support rate. This paper is giving public an idea of likely winner of the election that will have the potential implications on futrure political strategies and the overall direction of U.S. policy.
